Manuscript

### 2025-09-15 Manuscript_MD_ScoreReLiability.qmd
update manuscript to estimates of heterogeneity

For each replication of a phenomenon, Cohen’s $d$ is computed as an
estimate of raw ES, according to @eq-d, including its standard error
(@eq-SEd). Subsequently, a random-effects meta-analysis is performed
using metafor version 4.8-0 [@metafor]. To estimate meta-analytic mean
ES and heterogeneity we make use of the REML-estimator, as that
estimator tends to perform best across a range of conditions
[@hoenekopp2022]. To identify which phenomena pass criterion (3) we use
a Wald-type significance test we assess for which phenomenon the
meta-analytic mean ES is significantly different from zero. For this
hypothesis test, as all other hypothesis tests in this manuscript, we
employ a one-tailed significance level of .05.

Additionally, estimates of score reliability are derived. While not
without criticism (references), Cronbach’s Alpha is used to estimate
score reliability. It is often noted that tau-equivalence is an
unrealistic assumption to hold against real-world data, implying that
other estimates of score reliability, such as McDonald’s Omega
[@mcdonald1999], Guttman’s Lambda 2 or 4 [@guttman1945], or the Greatest
Lower Bound [@tenBerge2004] might be better suited. However, all scales
employed in this project are scored by computing the simple mean or sum
of responses across items for each individual. Computing the mean or sum
of several items implicitly assumes tau-equivalence, as each item
contributes equally to the individual’s test score. Therefore,
Cronbach’s Alpha is the better choice to estimate score reliability, as
it avoids a mismatch in assumptions between how the score is computed
and how score reliability is estimated. The individual estimates of
Cohen’s d are corrected using these estimates of Cronbach's Alpha,
according to @eq-dc, including the corrected standard errors (@eq-SEdc).
Thereby, estimates of ES are computed, which are corrected for imperfect
reliability. Again, a random-effects meta-analysis is run using metafor,
generating an estimate of heterogeneity in corrected ES. Different
indicators of heterogeneity are readily available, such as $I^2$, $H^2$
or the coefficient of variation $CV$[@borenstein2009; @higgins2002].
However, the descriptions found in H&S and W&D discuss the absolute
amount of heterogeneity in terms of variance ($\tau^2$) or standard
deviation ($\tau$). These parameters discuss the variability of the
standardized ES in the population in terms of variance or standard
deviation. Therefore, we also make use of absolute heterogeneity here.

Subsequently, in order to assess whether ES heterogeneity was indeed
reduced by the attenuation correction procedure, the estimates of
heterogeneity in uncorrected ES and corrected ES are compared. Using
Cochran’s *Q*-test [@cochran1954], we assess whether the estimates of
heterogeneity are statistically significantly different from 0. Only for
those projects where we have sufficient confidence that the individual
estimates of heterogeneity are larger than zero can we actually begin to
interpret whether there was any change.

To identify whether a reduction in ES heterogeneity is indeed
accompanied by differences in score reliability, the estimates of score
reliability are assessed meta-analytically as well. To do so, a
Reliability Generalization Meta-Analysis is performed [@vacha1998;
@sanchez2013]. This entails that individual estimates of score
reliability are adequately transformed using Bonett’s transformation:
$T_{\hat{\rho}} = ln(1 - \hat{\rho})$. This transformation has
variance-stabilising properties and therefore allows for adequate
inferences concerning heterogeneity in score reliability [@bonett2002].
Cochran’s Q-test is used to identify statistically significant
heterogeneity in transformed score reliability. Estimates derived from a
Reliability Generalization Meta-Analysis using these transformations can
be back-transformed to the original score reliability scale
[@beinhauer2025]:

@@

[Equations @eq-d], [-@eq-SEd], [-@eq-dc], and [-@eq-SEdc] were used to
generate estimates of standardized ES, both uncorrected and corrected,
with their corresponding standard errors. Generally, this leads to
larger (absolute) ES and larger standard errors after the correction. As
an example from the 21 selected phenomena, the results from this
procedure on replications on Nosek’s study on sex differences in
preferences for the subject of Arts is displayed in the forest plot in
@fig-forest. Here, grey dots represent the uncorrected ES for each
sample, while black dots represent the corrected ES in each sample. The
bars surrounding these dots show the respective 95%-Cis.

```{r, echo = FALSE, fig.height = 6}
#| label: fig-forest 
#| fig-cap: Forest Plot ML1 - Nosek (Art)
forest_plot_rel(ES_rma_df[which(ES_rma_df$MASC == "Nosek_Explicit_Art"),], agg_L[[which(MASC_names == "Nosek_Explicit_Art")]])
```

In @fig-forest it becomes apparent that the reliability attenuation
procedure leads to an increase in the individual absolute effect size,
as the black dots are moved further away from zero. Simultaneously, the
standard errors grew larger after the attenuation correction, which
leads to larger 95%-CIs. This is most easily observed for the
mturk-sample, which already had a rather large confidence interval to
begin with. Additionally, in the diamond at the bottom of the figure, we
can see that the meta-analytic estimate mean ES is also larger after the
attenuation correction took place. In @tbl-meanES, the estimates and
tests on the meta-analytic average ES, both corrected and uncorrected,
can be found.

@@

gt::gt(read.csv(here("Tables/Mean_ES.csv")) %>% 
         arrange(MASC)) %>% 
  gt::cols_label(MASC = "MASC",
                 mu_str_raw = gt::md("$\\hat{\\mu}_{\\delta_0}$ (SE)"),
                 pval_raw = "p",
                 mu_str_cor = gt::md("$\\hat{\\mu}_{\\delta}$ (SE)"),
                 pval_cor = "p") %>% 
  gt::tab_options(table.width = gt::pct(100),
                  column_labels.font.weight = "bold") # %>% 
  # gt::cols_width(pval_raw ~ gt::px(40),
  #                pval_cor ~ gt::px(40))
```

*Note:* Since in a meta-analysis, we attempt to estimate mean and
variance of the parameter's true distribution, in this table
$\hat{\mu}_{\delta_0}$ represents the meta-analytic estimate of mean
uncorrected $d$ and $\hat{\mu}_{\delta}$ represents the estimate of the
meta-analytic mean of corrected $d_c$.

>>>>

For each replication of a phenomenon, Cohen’s $d$ is computed as an
estimate of raw ES, according to @eq-d, including its standard error
(@eq-SEd). Subsequently, a random-effects meta-analysis is performed
using metafor version 4.8-0 [@metafor]. To estimate meta-analytic mean
ES and heterogeneity we make use of the REML-estimator, as that
estimator tends to perform best across a range of conditions
[@hoenekopp2022]. To identify which phenomena pass criterion (3) we use
a Wald-type significance test, assessing for which phenomenon the
meta-analytic mean ES is significantly different from zero. For this
hypothesis test, as all other hypothesis tests in this manuscript, we
employ a one-tailed significance level of .05.

Additionally, estimates of score reliability are derived. While not
without criticism [e.g. @sijtsma2009], Cronbach’s Alpha is used to
estimate score reliability. It is often noted that tau-equivalence is an
unrealistic assumption to hold against real-world data, implying that
other estimates of score reliability, such as McDonald’s Omega
[@mcdonald1999], Guttman’s Lambda 2 or 4 [@guttman1945], or the Greatest
Lower Bound [@tenBerge2004] might be better suited. However, all scales
employed in this project are scored by computing the simple mean or sum
of responses across items for each individual. Computing the mean or sum
of several items implicitly assumes tau-equivalence, as each item
contributes equally to the individual’s test score. Therefore,
Cronbach’s Alpha is the better choice to estimate score reliability, as
it avoids a mismatch in assumptions between how the score is computed
and how score reliability is estimated. The individual estimates of
Cohen’s $d$ are corrected using estimates of Cronbach's Alpha, according
to @eq-dc, including the corrected standard errors (@eq-SEdc). Thereby,
estimates of ES are computed, which are corrected for imperfect
reliability. Again, a random-effects meta-analysis is run using metafor,
generating an estimate of heterogeneity in corrected ES. Different
indicators of heterogeneity are readily available, such as $I^2$, $H^2$
or the coefficient of variation $CV$[@borenstein2009; @higgins2002].
However, the descriptions found in H&S and W&D discuss the absolute
amount of heterogeneity in terms of variance ($\tau^2$) or standard
deviation ($\tau$). Therefore, we also make use of absolute
heterogeneity here.

Subsequently, in order to assess whether ES heterogeneity was indeed
reduced by the attenuation correction procedure, the estimates of
heterogeneity in uncorrected ES and corrected ES are compared. Using
Cochran’s *Q*-test [@cochran1954], we assess whether the estimates of
heterogeneity are statistically significantly different from 0. Only for
those projects where we have sufficient confidence that the individual
estimates of heterogeneity are larger than zero can we actually begin to
interpret whether there was any change. \[Short paragraph discussing
that a test for difference in heterogeneity would be more informative?
Maybe move it to the discussion section - hint at it here?\]

To identify whether a reduction in ES heterogeneity is indeed
accompanied by differences in score reliability, the estimates of score
reliability are assessed meta-analytically as well. To do so, a
Reliability Generalization Meta-Analysis is performed [@vacha1998;
@sanchez2013]. This entails that individual estimates of score
reliability are adequately transformed using Bonett’s transformation:

$$T_{\hat{\rho}} = ln(1 - \hat{\rho})$$ {#eq-Bonett}

The transformation in @eq-Bonett has variance-stabilising properties and
therefore allows for adequate inferences concerning heterogeneity in
score reliability [@bonett2002]. Cochran’s Q-test is used to identify
statistically significant heterogeneity in transformed score
reliability. Estimates derived from a Reliability Generalization
Meta-Analysis using these transformations can be back-transformed to the
original score reliability scale [@beinhauer2025]:

@@

[Equations @eq-d], [-@eq-SEd], [-@eq-dc], and [-@eq-SEdc] were used to
generate estimates of standardized ES, both uncorrected and corrected,
with their corresponding standard errors. Generally, this leads to
larger (absolute) ES and larger standard errors after the correction. As
an example from the 21 selected phenomena, the results from this
procedure on replications of Nosek’s study on sex differences in
preferences for the subject of Arts is displayed in the forest plot in
@fig-forest. Here, grey dots represent the uncorrected ES for each
sample, while black dots represent the corrected ES in each sample. The
bars surrounding these dots show the respective 95%-Cis.

```{r, echo = FALSE, fig.height = 6}
#| label: fig-forest 
#| fig-cap: Forest Plot ML1 - replications of Nosek's studz on sex differences in subject preferences (Arts)
forest_plot_rel(ES_rma_df[which(ES_rma_df$MASC == "Nosek_Explicit_Art"),], agg_L[[which(MASC_names == "Nosek_Explicit_Art")]])
```

In @fig-forest it becomes apparent that the reliability attenuation
procedure leads to an increase in the individual absolute effect size,
as the black dots are moved further from zero. Simultaneously, the
standard errors grew larger after the attenuation correction, which
leads to larger 95%-CIs. This is most easily observed for the
MTurk-sample, which had a rather large confidence interval to begin
with. Additionally, in the diamond at the bottom of the figure, we can
see that the meta-analytic mean ES is also larger after the attenuation
correction took place. @tbl-meanES contains estimates and test results
for all phenomena, both corrected and uncorrected.

@@

gt::gt(read.csv(here("Tables/Mean_ES.csv")) %>% 
         arrange(MASC)) %>% 
  gt::cols_label(MASC = "MASC",
                 #mu_str_raw = gt::md("$\\hat{\\mu}_{\\delta_0}$ (SE)"),
                 mu_str_raw = gt::md("Mean $d$ (SE)"),
                 pval_raw = "p",
                 #mu_str_cor = gt::md("$\\hat{\\mu}_{\\delta}$ (SE)"),
                 mu_str_cor = gt::md("Mean $d_c$ (SE)"),
                 pval_cor = "p") %>% 
  gt::tab_options(table.width = gt::pct(100),
                  column_labels.font.weight = "bold") # %>% 
  # gt::cols_width(pval_raw ~ gt::px(40),
  #                pval_cor ~ gt::px(40))
```

*Note:* In this table *Mean* $d$ represents the meta-analytic estimate
of mean uncorrected $d$ and *Mean* $d_c$ represents the estimate of the
meta-analytic mean of corrected $d_c$.
